---
title: "Commodities Analysis"
author: "Mariam Iftikhar"
date: "2025-06-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# ðŸ“¥ Load Libraries
library(quantmod)
library(TTR)
```

```{r}
# ðŸ“Š Choose your commodity ETF
symbol <- "PPLT"  # Change to "SLV", "GLD", "USO", etc.

# ðŸ•° Get historical data
getSymbols(symbol, from = "2022-01-01", auto.assign = TRUE)
data <- get(symbol)

# ðŸ“ˆ Technical Indicators
price <- Cl(data)
rsi <- RSI(price, n = 14)
macd <- MACD(price, nFast = 12, nSlow = 26, nSig = 9)
sma50 <- SMA(price, n = 50)
sma200 <- SMA(price, n = 200)

# ðŸ§  Entry/Exit Strategy Logic
signals <- data.frame(Date = index(data),
                      Price = as.numeric(price),
                      RSI = as.numeric(rsi),
                      MACD = as.numeric(macd$macd),
                      Signal = as.numeric(macd$signal),
                      SMA50 = as.numeric(sma50),
                      SMA200 = as.numeric(sma200),
                      stringsAsFactors = FALSE)
summary(signals)


# Re-initialize Position column
signals$Position <- NA

for (i in 2:nrow(signals)) {
  if (!is.na(signals$RSI[i]) &&
      !is.na(signals$SMA50[i]) &&
      !is.na(signals$SMA200[i]) &&
      !is.na(signals$MACD[i]) &&
      !is.na(signals$Signal[i]) &&
      !is.na(signals$MACD[i-1]) &&
      !is.na(signals$Signal[i-1])) {
    
    if (signals$RSI[i] < 30 && signals$SMA50[i] > signals$SMA200[i]) {
      signals$Position[i] <- "BUY"
    } else if (signals$RSI[i] > 70 || 
               (signals$MACD[i] < signals$Signal[i] && signals$MACD[i-1] > signals$Signal[i-1])) {
      signals$Position[i] <- "SELL"
    } else {
      signals$Position[i] <- "HOLD"
    }
  } else {
    signals$Position[i] <- NA
  }
}

tail(signals[, c("Date", "Price", "RSI", "SMA50", "SMA200", "MACD", "Signal", "Position")])


# ðŸ–¨ï¸ View latest signal
tail(signals, 10)

# ðŸ“ˆ Optional: Plot with chart
chartSeries(data, TA = c(addSMA(50), addSMA(200), addRSI(), addMACD()))
```

## Explanation of Code:

1. quantmod is a popular package to get financial data and create charts.

2. TTR provides technical indicators like RSI, MACD, and moving averages.

3. symbol holds the ticker of the commodity ETF you want to analyze.

4. getSymbols() downloads historical price data from Yahoo Finance starting from Jan 1, 2022.

5. data stores the downloaded data (daily OHLCV prices).

6. RSI: Measures momentum; values below 30 may indicate oversold (buy signal), above 70 overbought (sell signal).

7. MACD: Measures trend and momentum; the difference between two EMAs (fast=12, slow=26) plus a signal line (9).

8. SMA50 and SMA200: Trend indicators; 50-day SMA crossing above 200-day SMA usually indicates a bullish trend.

9. Combines the date, price, and all indicators into one table (signals).

10. Converts all data to numeric for easier calculations.

11. This makes it easy to analyze day-by-day and generate trading signals.

12. Creates a new column Position to store the trading decision for each day: "BUY", "SELL", or "HOLD".

13. The loop iterates over each day from the second row onward.

14. It only processes days where none of the required indicator values are NA (because early days may have missing SMA, MACD, or RSI).

15. BUY signal:

  a. RSI < 30 (oversold, likely price rebound)

  b. 50-day SMA is above 200-day SMA (indicating a bullish trend)

16. SELL signal:

  a. RSI > 70 (overbought, possible price drop)

  b. OR MACD line crosses below the signal line (bearish crossover)

17. HOLD signal means no clear buy or sell conditions met.

18. If insufficient data (NA in any indicators), position is set to NA.
The loop iterates over each day from the second row onward.

19. It only processes days where none of the required indicator values are NA (because early days may have missing SMA, MACD, or RSI).

20. BUY signal:

  a. RSI < 30 (oversold, likely price rebound)

  b. 50-day SMA is above 200-day SMA (indicating a bullish trend)

21. SELL signal:

  a. RSI > 70 (overbought, possible price drop)

  b. OR MACD line crosses below the signal line (bearish crossover)

  c. HOLD signal means no clear buy or sell conditions met.

  d. If insufficient data (NA in any indicators), position is set to NA.

22. Uses quantmod's chartSeries to plot the ETF price chart.

23. Adds 50-day and 200-day SMAs as lines on the price chart.

24. Adds RSI and MACD plots below the price chart for momentum and trend visualization.

25. This visual helps confirm signals and analyze trends.